%-*- TeX -*- -*- Soft -*-,

\documentclass[11pt]{bmc_article_s50}

\usepackage{amsthm,amsmath}
\usepackage{siunitx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[utf8]{inputenc}
\RequirePackage{hyperref}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{url}
\urlstyle{rm}
\usepackage{setspace}
\usepackage[T1]{fontenc}
\pagestyle{empty}
\setlength{\parindent}{0cm}
\usepackage{ragged2e}
\justifying
\singlespace

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
  
\begin{document}

\title{Brainhack: A new kind of conference for a new kind of scientific
community}

\maketitle

\author[1,2*]{R Cameron Craddock}\cor{}\email{ccraddock@nki.rfmh.org}
\author[3]{Daniel S Margulies}\email{margulies@cbs.mpg.de}


\address[1]{
  Computational Neuroimaging Lab, Center for Biomedical Imaging and
Neuromodulation, Nathan Kline Institute for Psychiatric Research%
  , 140 Old Orangeburg Rd%
  , 10962%
  , Orangeburg%
  , New York %
  , USA
}
\address[2]{
  Center for the Developing Brain, Child Mind Institute%
  , 445 Park Ave%
  , 10022%
  , New York%
  , New York %
  , USA
}
\address[3]{
  Max Planck Research Group for Neuroanatomy and Connectivity, Max Planck
Institute for Human Cognitive and Brain Sciences%
  , Stephaßnstraße 1A%
  , 4103%
  , Leipzig%
   %
  , Germany
}


\begin{abstract}
\emph{Should be no more than 100 words}
\end{abstract}


\keywords{hackathon, unconference, open science, neuroscience, data sharing,
collaboration}

\emph{Brainhack Commentary}
\emph{http://www.gigasciencejournal.com/authors/instructions/commentary}

\section{Brainhack}\label{brainhack}

Now in their third year, international and regional Brainhack events
bring together brain enthusiasts from a variety of backgrounds to build
relationships, learn from one another, and work on projects related to
the brain. Brainhacks are unconventional events that adapt concepts from
hackathons and unconferences to brain research. Similar to hackathons,
Brainhack events primarily involve unstructured time during which
participants collaborate intensitvely to complete various projects.
Rather than focusing on computer programming, projects at the Brainhack
can be completed using a much broader array of methods. Time is also set
aside for periodic ``unconference'' sessions, which involve
presentations whose content is dynamicaly determined based on the
interests of attendees. This combined model encourages active
participation and interaction between attendees and maximizes the
relevance of the content.

There is no ideal background, skillset or experience level for Brainhack
attendees. Fully translating neuroscience data to knowledge requires
expertise that spans the gamut from biology, computer science,
engineering, informatics, mathematics, neuroanatomy, philosopy, physics,
psychiatry, psychology, statistics, art, and many others. The goal of
the Brainhack is to facilitate the cross pollenation of ideas and
knowledge across these various disciplines and communities, to
accelerate the development of a richer understanding of the brain. In
addition to sharing data and tools, attendees can contribute to this
agenda in a variety of ways. Philosophical debates about the meanings of
cognition, coordinated efforts to manually segment brain images from
different species, curating neuroscience literature, or helping others
understand the subtelties of diagnosing a developmental disorder are all
examples of extremely valuable contributions to the Brainhack.

\section{Hackathons based on collaboration, not
competition}\label{hackathons-based-on-collaboration-not-competition}

The hackathon format gained prominence in the technology sector by
providing a meeting model that targets specific project goals during
intense time-limited collaborations. The competitive aspect of the
traditional hackathon, while catalyzing advances toward specific
techology ends, is contrary to the founding motive of Brainhack, which
is to encourage open, cross-institutional and inter-disciplinary
collaboration. Rather than subdividing attendees into competetive
factions, each of whom are trying to solve the exact same prespecified
problem, the ``Brainhack way'' is to encourage attendees to work
together in collaborative teams to solve problems of their choosing. In
this way we have not only attained solutions to many problems, rather
than many solutions to a single problem, but we have pollenated new
relationships that will hopefully be productive well into the future.


\begin{table}[!h]
\caption{{\bf Brainhack events occuring 2012 - 2013.} The number of attendees for each event are included in parenthesis. $^\star$Local organizers for an event. $^\dagger$Organizer for a distributed event.}\label{tab1}
  \begin{tabular}{L{.1in}L{.1in}L{2.4in}L{2in}L{1.5in}}
    \hline
	\\
    \multicolumn{3}{l}{\textbf{2012 Brainhack and Unconference}} &  & September 1 - 4, 2013  \\
	& \multicolumn{3}{l}{Ann Arbor, Michigan, USA (77)} \\
    & & \multicolumn{3}{l}{Host: Max Planck Institute for Human Cognition and Brain Sciences} \\
    & & \multicolumn{3}{l}{Organizers: Daniel Margulies$^{\star}$, Pierre Bellec, Cameron Craddock, Donald McLaren,} \\
    & & \multicolumn{3}{l}{\hspace*{25pt}Maarten Mennes}\\
	\\
    \multicolumn{3}{l}{\textbf{Brainhack 2013}} & & October 23 - 26, 2013  \\
	& \multicolumn{3}{l}{Paris, France (77)} \\
    & & \multicolumn{3}{l}{Host: Laboratoire d'Imagerie Biom\'{e}dicale, Sorbonne Universit\'{e}s, Universit\'{e} Pierre-et- } \\
    & & \multicolumn{3}{l}{\hspace*{25pt}Marie-Curie, Paris 06, CNRS, INSERM} \\
    & & \multicolumn{3}{l}{Organizers: Selma Mesmoudi$^{\star}$, Yves Burnod$^{\star}$, Donald McLaren, Cameron Craddock,} \\
    & & \multicolumn{3}{l}{\hspace*{25pt}Pierre Bellec, Daniel Margulies, Maarten Mennes} \\
	\\
    \multicolumn{3}{l}{\textbf{OHBM Hackathon 2014}} & & July 5 - 7, 2014  \\
	& \multicolumn{3}{l}{Berlin, Germany (77)} \\
    & & \multicolumn{3}{l}{Host: Organization for Human Brain Mapping } \\
    & & \multicolumn{3}{l}{Organizers: Daniel Margulies$^{\star}$, Pierre Bellec, Cameron Craddock, Tom Grabowski,} \\
    & & \multicolumn{3}{l}{\hspace*{25pt}Sean Hill, Nolan Nichols, JB Poline} \\
	\\
    \multicolumn{3}{l}{\textbf{Brainhack Eastern Daylight Time}} & & October 18 - 19, 2014  \\
    & \multicolumn{3}{l}{Ann Arbor, Michigan, USA (77)} \\
    & & \multicolumn{3}{l}{Host: University of Michigan } \\
    & & \multicolumn{3}{l}{Organizers: Scott Peltier$^{\star}$, Robert Welsh$^{\star}$} \\
    & \multicolumn{3}{l}{Boston, Massachusetts, USA (77)} \\
    & & \multicolumn{3}{l}{Host: Massachusetts Institute of Technology} \\
    & & \multicolumn{3}{l}{Organizers: Satra Gosh$^{\star}$, Matt Hutchison$^{\star}$, Donald McLaren$^{\star}$} \\
    & \multicolumn{3}{l}{Miami, Florida, USA (77)} \\
    & & \multicolumn{3}{l}{Host: Florida International University} \\
    & & \multicolumn{3}{l}{Organizers: Angie Laird$^{\star}$, Lucina Uddin$^{\star}$} \\
    & \multicolumn{3}{l}{New York, New York, USA (77)} \\
    & & \multicolumn{3}{l}{Host: Child Mind Institute and Columbia University} \\
    & & \multicolumn{3}{l}{Organizers: Cameron Craddock$^{\star\dagger}$, Andrew Gerber$^{\star}$} \\
    & \multicolumn{3}{l}{Porto Alegre, Brazil (77)} \\
    & & \multicolumn{3}{l}{Host: Pontif\'{i}cia Universidade Cat\'{o}lica do Rio Grande do Sul} \\
    & & \multicolumn{3}{l}{Organizers: Alex Franco$^{\star}$, Caroline Fr\"{o}hlich$^{\star}$, Felipe Meneguzzi$^{\star}$} \\
    & \multicolumn{3}{l}{Toronto, Ontario, Canada (77)} \\
    & & \multicolumn{3}{l}{Host: University of Toronto} \\
    & & \multicolumn{3}{l}{Organizers: Jonathan Downer$^{\star}$, Katie Dunlop $^{\star}$, Stephen Strother$^{\star}$} \\
    & \multicolumn{3}{l}{Washington DC, USA (77)} \\
    & & \multicolumn{3}{l}{Host: Georgetown University} \\
    & & \multicolumn{3}{l}{Organizers: John Van Meter$^{\star}$, Lei Liew$^{\star}$, Ziad Saad$^{\star}$, Prantik Kundu$^{\star}$} \\
	\\
	\hline
	\end{tabular}
\end{table}

\begin{table}[!h]
\caption{{\bf Brainhack events in 2015.} The number of attendees for each event are included in parenthesis. $^\star$Local organizers for an event. $^\dagger$Organizer for a distributed event.}\label{tab2}
  \begin{tabular}{L{.1in}L{.1in}L{2.4in}L{2in}L{1.5in}}
    \hline
	\\
    \multicolumn{3}{l}{\textbf{OHBM Hackathon 2015}} & & June 12 - 14, 2015  \\
	& \multicolumn{3}{l}{Honolulu, Hawaii, USA (77)} \\
    & & \multicolumn{3}{l}{Host: Organization for Human Brain Mapping } \\
    & & \multicolumn{3}{l}{Organizers: Nolan Nichols$^{\star}$, Pierre Bellec, Cameron Craddock, Tom Grabowski,} \\
    & & \multicolumn{3}{l}{\hspace*{25pt}Jack Van Horn, Daniel Margulies, JB Poline} \\
	\\
    \multicolumn{3}{l}{\textbf{Brainhack Montreal 2015}} & & July 27 - 29, 2015 \\
	& \multicolumn{3}{l}{Montr\'{e}al, Queb\'{e}c, Canada (77)} \\
    & & \multicolumn{3}{l}{Host: Centre de recherche de l'Institut universitaire de g\'{e}riatrie de Montr\'{e}al} \\
    & & \multicolumn{3}{l}{Organizers: Benjamin De Leener$^{\star}$, Julien Cohen-Adad$^{\star}$, Pierre Bellec$^{\star}$} \\
    & & \multicolumn{3}{l}{\hspace*{25pt}Sebastien Dery$^{\star}$, Pierre-Olivier Quirion$^{\star}$} \\
	\\
    \multicolumn{3}{l}{\textbf{Brainhack Americas}} & & October 23 - 25, 2015  \\
    & \multicolumn{3}{l}{Ann Arbor, Michigan, USA (77)} \\
    & & \multicolumn{3}{l}{Host: University of Michigan } \\
    & & \multicolumn{3}{l}{Organizers: Scott Peltier$^{\star}$, Robert Welsh$^{\star}$} \\
	
    & \multicolumn{3}{l}{Berkeley, California, USA (77)} \\
    & & \multicolumn{3}{l}{Host: D-Lab University of California, Berkeley} \\
    & & \multicolumn{3}{l}{Organizers: Daniel Lurie$^{\star}$, JB Poline$^{\star}$} \\
	
    & \multicolumn{3}{l}{Los Angeles, California, USA (77)} \\
    & & \multicolumn{3}{l}{Host: University of Southern California} \\
    & & \multicolumn{3}{l}{Organizers: Lei Liew$^{\star}$, Gautam Prasad$^{\star}$, Yonggang Shi$^{\star}$} \\
	
    & \multicolumn{3}{l}{Miami, Florida, USA (77)} \\
    & & \multicolumn{3}{l}{Host: University of Miami} \\
    & & \multicolumn{3}{l}{Organizers: Angie Laird$^{\star}$, Lucina Uddin$^{\star}$} \\
	
    & \multicolumn{3}{l}{New York, New York, USA (77)} \\
    & & \multicolumn{3}{l}{Host: Translational \& Molecular Imaging Institute, Icahn School of Medicine} \\
	& & \multicolumn{3}{l}{\hspace*{25pt}at Mount Sinai and Child Mind Institute} \\
    & & \multicolumn{3}{l}{Organizers: Cameron Craddock$^{\dagger}$, Christopher Cannistraci$^{\star\dagger}$, Prantik Kundu$^{\star\dagger}$,} \\
	& & \multicolumn{3}{l}{\hspace*{25pt}David O'Connor, Ting Xu} \\
	
    & \multicolumn{3}{l}{Porto Alegre, Brazil (77)} \\
    & & \multicolumn{3}{l}{Host: Pontif\'{i}cia Universidade Cat\'{o}lica do Rio Grande do Sul} \\
    & & \multicolumn{3}{l}{Organizers: Alex Franco$^{\star}$, Felipe Meneguzzi$^{\star}$} \\
	
    & \multicolumn{3}{l}{Quer\'{e}taro, M\'{e}xico (77)} \\
    & & \multicolumn{3}{l}{Host: Instituto De Neurobiolog\'{i}a, Universidad Nacional Aut\'{o}noma de M\'{e}xico} \\
    & & \multicolumn{3}{l}{Organizers: Sarael Alcauter$^{\star}$, Fernando Barrios$^{\star}$, Cameron Craddock$^{\dagger}$,}\\
	& & \multicolumn{3}{l}{\hspace*{25pt}Eric H Pasaye$^{\star}$} \\
	
    & \multicolumn{3}{l}{Seattle, Washington, USA (77)} \\
    & & \multicolumn{3}{l}{Host: University of Washington eScience Institute} \\
    & & \multicolumn{3}{l}{Organizers: Ariel Rokem$^{\star}$} \\
	\\
	\hline
	\end{tabular}
\end{table}

Previous Brainhack events (see Table XX) have spanned one to three days
and include a variety of content to make them accessible and fruitful
for a wide audience of attendees. The format of the events is not fixed,
but varies based on what the local organizing committee deams most
salient to attendees. In general, a few different types of content are
available at Brainhack events:

\begin{itemize}
\tightlist
\item
  \textbf{Meet and great:} Brainhack events begin with a welcome to the
  hosting facility by the local organizing committee, along with a
  briefing about the event schedule, procedures, or other information
  that might be important for the attendees.
\item
  \textbf{Ice breaker:} Interaction between attendees is the key to a
  successful Brainhack event. The ice breaker is an activity to
  introduce attendees and their interestes on one another. One strategy
  that has been successful is for each attendee, in turn, give their
  name, their institution, and three words that describe their
  interests. We have toyed with the idea of a speed-dating paradigm in
  which attendees pair-up for a brief conversation, after 2 minutes
  attendees swap partners, and this continues until every pair of
  attendees has met. This strategy has yet to be implemented at any of
  the Brainhack events.
\item
  \textbf{Ignite talks:} Brainhack's equivalent of keynote sessions,
  Ignite talks are inspirational talks on the big picture of open brain
  science that are intended to invigorate the audience and set the
  agenda for the day. These are ideally short, usually 20 minutes with
  10 minutes of questions, and are provided by a luminary in the field.
\item
  \textbf{Hacking:} The core of Brainhack is ``open hacking'' sessions
  during which attendess collaborate together on projects of their
  choosing. Attendees who have specific project ideas or data that they
  would like to explore are encouraged to advertise their project at
  \href{http://www.brainhack.org}{www.brainhack.org} prior to the event.
  On the opening of the Brianhack event, typically after the ice
  breaker, attendees pitch their ideas and afterwards mingle with others
  to organize a project team. Teams work together throughout the
  remainder of Brainhack and are given the opportunity to present their
  progress during the wrap-up session at the end of the event.
\item
  \textbf{Brainhacking 101:} In response to feedback from previous
  Brainhack events, an educational track has been added to teach basic
  software and data analysis skills to researchers who are weak in these
  areas. The educational track occurs in parallel, and is designed not
  to interfere, with the hacking track. The track begins with
  ``Installfest'' sessions during which attendees receive help
  installing needed software. Afterwards are several hands-on tutorials
  that cover topics like: using Github, Python programming, using Python
  to load and visualize neuroimaging data, and performing meta-analyses
  of scientific literature.
\item
  \textbf{Unconference:} Time is set aside for attendees to present on
  their research or other topics of interest. The agenda for these
  sessions are determined onsite an hour or two before they are to
  occur. Attendees who are willing to present add their name to a signup
  sheet and in the event that there are more interested presenters than
  time, the group is polled to determine which presentations are given
  time or to extend the amount of time alloted. Instead of unconference
  sessions, some sites have incorporate ``Data Blitzes'', which is a
  pre-organized session that provides interested attendees the
  opportunity to present on their research. Brainhack Miami has had a
  lot of success with this model and has secured funding to award
  monetary prizes to the best presentations.
\item
  \textbf{Wrapup and feedback:} Brainahack events typically end with a
  wrap-up session during which project teams describe the progress that
  they made or give a demo of their results. Afterwards, the local
  organizers lead a discussion about what worked well with the event,
  and what should be changed in the future.
\end{itemize}

\subsection{Project examples}\label{project-examples}

Following are examples of projects that have previously been initiated
and/or completed at Brainhack evens:

\begin{itemize}
\item
  \cite{Mennes2014}
  \href{http://www.sciencedirect.com/science/article/pii/S1053811914002973}{Mennes
  et al 2014} pooled data at brainhack to reach a sample size of
  N=14,781 in their imaging study of optimization of field of view for
  maximal brain coverage.
\item
  \href{http://www.sciencedirect.com/science/article/pii/S1053811915006898}{Vanderwal
  et al In Press} collaboration between a 3D artist and a child
  psychiatrist to reduce in-scanner motion in resting-state fMRI using a
  movie specifically tailored for that purpose.
  \href{http://vimeo.com/67962604}{video link}
\item
  \href{http://journal.frontiersin.org/article/10.3389/fninf.2011.00037/abstract}{Das
  et al. 2012} the development team of LORIS, an open-source database
  system for neuroimaging and phenotypic data, have repeatedly used
  brainhack as an opportunity to meet and collaborate on new features.
\item
  Toro's \href{http://braincatalogue.org}{braincatalogue.org}
\item
  \href{http://daydreaming-the-app.net}{Daydreaming app}
\item
  \href{http://onlinelibrary.wiley.com/doi/10.1002/hbm.22776/full}{Multi-center
  mapping of structural network alterations in autism}
\item
  The is a consortium of neuroimaging reserchers who are preprocessing
  the openly shared
  \href{http://fcon_1000.projects.nitrc.org/indi/abide/}{Autism Brain
  Imaging Data Exchange (ABIDE)} data using the
  \href{http://www.google.com}{Connectome Computation System (CCS)},
  \href{http://fcp-indi.github,io}{Configurable Pipeline for the
  Analysis of Connectomes}, \href{http://www.google.com}{Data
  Preprocessing Assistant for Resting State fMRI (DPARSF)},
  \href{http://www.google.com}{Neuro Imaging Analysis Kit (NIAK)},
\item
  The
  \href{http://preprocessed-connectomes-project.github.io/abide}{ABIDE
  Preprocessing Initiative} is an ongoing project that started at the
  2012 Brainhack and UnConference in Leipzig to share preprocessed
  versions of the
  \href{http://fcon_1000.projects.nitrc.org/indi/abide}{Autism Brain
  Imaging Data Exchange (ABIDE)} dataset. This project is sharing
  functional data that have been processed using the
  \href{http://lfcd.psych.ac.cn/ccs.html}{Connectome Computation System
  (CCS)}, the \href{http://fcp-indi.github.io}{Configurable Pipeline for
  the Analysis of Connectomes (C-PAC)}, the
  \href{http://rfmri.org/DPARSF}{Data Preprocessing Assistant for
  Resting State fMRI (DPARSF)}, and the
  \href{https://www.nitrc.org/projects/niak/}{Neuro Imaging Analysis Kit
  (NIAK)}. Cortical thickness measures have been extracted from
  structural data using \href{http://freesurfer.net/}{FreeSurfer},
  \href{http://www.bic.mni.mcgill.ca/ServicesSoftware/CIVET}{CIVET}, and
  \href{http://picsl.upenn.edu/software/ants/}{Advanced nromalization
  Tools (ANTS)}.
\end{itemize}

\subsection{Future challenges}\label{future-challenges}

\begin{itemize}
\tightlist
\item
  series of events that culminate at the OHBM hackathon??
\item
  variation of the format into a teaching-oriented event?
\end{itemize}

\subsubsection{Distributed events}\label{distributed-events}

\begin{itemize}
\tightlist
\item
  describe types of brainhack events
\end{itemize}

\subsubsection{Hosting a Brainhack}\label{hosting-a-brainhack}

\begin{itemize}
\tightlist
\item
  encourage others to host brainhacks in their own areas
\end{itemize}

\subsubsection{Post-conference
publications}\label{post-conference-publications}

\begin{itemize}
\tightlist
\item
  motivate post-conference publications and brainhack thematic series
\end{itemize}

\section{Conclusions}\label{conclusions}

%\theendnotes

\section*{Availability of Supporting Data}
More information about this project can be found at: \url{http://github.com/ccraddock/afni}. Further data and files supporting this project are hosted in the \emph{GigaScience} repository REFXXX.

\section*{Competing interests}
None

\section*{Author's contributions}
RCC and DJC wrote the software, DJC performed tests, and DJC and RCC
wrote the report.

\section*{Acknowledgements}
The authors would like to thank the organizers and attendees of
Brainhack MX and the developers of AFNI. This project was funded in part
by a Educational Research Grant from Amazon Web Services.

  
  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file
\bibliography{brainhack-commentary} % Bibliography file (usually '*.bib' )

\end{document}
